---
title: "Indicateurs"
author: "FD"
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---


**NOTE** I am just using Rmd here to organize my code. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.par = TRUE)
```


```{r}
today <- Sys.Date() # today's date
```

```{r}
# Whether to generate a gif
doGif <- TRUE
lang <- "FR"#"EN" #"FR" 

dataVersion <- 2
# 1: indicateurs dataset
# 2: hospitalisation dataset

source("usefulFunctions.R")

```


Mise à jour du `r today`.   


# Load and clean data

## Version 1

```{r}
if(dataVersion == 1){
  # Source: <https://www.data.gouv.fr/fr/datasets/synthese-des-indicateurs-de-suivi-de-lepidemie-covid-19/>
  URLsource <- "https://www.data.gouv.fr/fr/datasets/synthese-des-indicateurs-de-suivi-de-lepidemie-covid-19/"
  URL <- "https://www.data.gouv.fr/fr/datasets/r/f335f9ea-86e3-4ffa-9684-93c009d5e617"
  dataFile <- paste0("data/FranceIndic_", today, ".csv") # name file with today's date
  download.file(URL, dataFile) # download file from repo
  dat.France <- read.csv(dataFile, sep = ",", stringsAsFactors = FALSE)
  
  # Remove leading NAs
  ibegin <- which(!is.na(dat.France$hosp | !is.na(dat.France$rea)))[1]
  dat.France <- dat.France[-(1:ibegin), ]
  
  # Compute sliding window average
  dat.France$incid_hosp_ave7 <- sliding.window(dat.France$incid_hosp, pos = 7)
  
}
```

## Version 2

TODO: add with/for columns

```{r, results='hide'}
if(dataVersion == 2){
  # Source: https://www.data.gouv.fr/fr/datasets/donnees-hospitalieres-relatives-a-lepidemie-de-covid-19/
  URLsource <- "https://www.data.gouv.fr/fr/datasets/donnees-hospitalieres-relatives-a-lepidemie-de-covid-19/"
  
  URL2 <- "https://www.data.gouv.fr/fr/datasets/r/fe3e7099-a975-4181-9fb5-2dd1b8f1b552"
  
  dataFile <- paste0("data/FranceHosp_", today, ".csv") # name file with today's date
  download.file(URL2, dataFile) # download file from repo
  dat.France2 <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)
  
  head(dat.France2)
  
  dat.France2.tot <- dat.France2[dat.France2$PourAvec == 0, ]
  dim(dat.France2.tot)
  
  # Rename columns
  # The data are per 100'000, 
  # and SPF used INSEE 2020 population estimations (given in the vaccine datasets)
  fact <- 67114995 / 10^5
  dat.France2.tot$incid_hosp_ave7 <- dat.France2.tot$tx_indic_7J_hosp * fact / 7
  dat.France2.tot$hosp <- dat.France2.tot$tx_prev_hosp * fact
  dat.France2.tot$rea <- dat.France2.tot$tx_prev_SC * fact
  
  # Sanity checks between the two versions
  # range(dat.France2.tot$hosp)
  # range(dat.France$hosp)
  
  # range(dat.France2.tot$incid_hosp_ave7 / 7)
  # range(dat.France$incid_hosp_ave7, na.rm = TRUE)
  
  # range(dat.France2.tot$rea)
  # range(dat.France$rea)
  
  dat.France <- dat.France2.tot
  dat.France$date <- dat.France$jour
}

```

## Clean dates

```{r eval = FALSE}
# Just for console evaluation
head(dat.France)
```


```{r}
# Transform dates as decimals
year <- substr(dat.France$date, 1, 4) # Extract year
nbdays <- rep(365, nrow(dat.France)) # Number of days in the year
nbdays[year == "2020"] <- 366 # 2020 is leap year

# Rewrite dates as fraction of the current year
dat.France$fracYear <- as.numeric(base::as.Date(dat.France$date) - base::as.Date(paste0(year, "-01-01"))) / nbdays # Fraction of the year

```

# Useful functions

```{r}
# Polar coordinates: r, theta
# x = r * cos(theta)
# y = r * sin(theta)

frac2angle <- function(frac){
  # Transform fraction of year into an angle in radian
  - frac * 2 * pi + pi/2
}

polar2xy <- function(r, frac){
  maxval <- max(abs(r), na.rm = TRUE) # Get max value to center the plot
  theta <- frac2angle(frac) # angle value in radian, rotate clockwise
  
  # Polar to Cartesian coordinates
  x <- r * cos(theta)
  y <- r * sin(theta)

  list(x = x, y = y, maxval = maxval)  
}

# col2021 <- "#0072B2"
# col2020 <- "#FF3900"#"#D55E00"
# col2022 <- "#FFBA00"

library("MetBrewer")

pal <- met.brewer("Egypt", n = 4, "discrete")

col2020 <- pal[2]
col2021 <- pal[4]
col2022 <- pal[3]



cols <- rep(col2020, nrow(dat.France))
cols[year == "2021"] <- col2021
cols[year == "2022"] <- col2022
makeTransparent<-function(someColor, alpha = 100){
  # Source : https://stackoverflow.com/questions/8047668/transparent-equivalent-of-given-color
  newColor <- col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red = curcoldata[1], 
                                              green = curcoldata[2],
                                              blue = curcoldata[3], 
                                              alpha = alpha, 
                                              maxColorValue = 255)})
}
colsTrp <- makeTransparent(cols, 70)
```

```{r}
addLegendPlot <- function(iend, graduations1, graduations2, r, dfinal, valfinal, yearfinal = 2022, addCircle = FALSE){
  # graduations1 main graduations for number in concentric circles
  # graduations2 sub graduations
  # r radius of months / time legend
  # dfinal fraction of the last day
  # valfinal value on the last day
  
  gmax <- max(c(graduations1, graduations2))
  lwd.grad <- 1.2
  
  for(i in graduations1){
    cxy <- polar2xy(i, seq(0, 1, length.out = 100))
    lines(cxy$x, cxy$y, type = "l", col = "gray", lwd = lwd.grad)
  }
  for(i in graduations2){
    cxy <- polar2xy(i, seq(0, 1, length.out = 100))
    lines(cxy$x, cxy$y, type = "l", col = "gray", lty = 3, lwd = lwd.grad)
  }
  for(i in graduations1){
  text(x = 0, y = graduations1, labels = graduations1, col = "gray", cex = 0.7, adj = c(0.5, -0.2) )
  text(x = 0, y = -graduations1, labels = graduations1, col = "gray", cex = 0.7, adj = c(0.5, +1.2) )
  }
  
  
  # Add months legend
  if(lang == "FR"){
    months <- c("Jan", "Fev", "Mar", "Avr", "Mai", "Jun", "Jui", "Aou", "Sep", "Oct", "Nov", "Dec")
  }else{
    months <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")    
  }
  
  thetas <- (1:12)/12 - 0.5/12 # Angles of the months
  for(i in seq_along(months)){
    text(x = r * cos(frac2angle(thetas[i])), y = r * sin(frac2angle(thetas[i])), labels = months[i], srt = 90 - thetas[i]*360, col = gray(0.5))
  }
  
  # Add line to locate the current days
  
  # Line to show the date
  lines(c(0, gmax * cos(dfinal)), c(0, gmax * sin(dfinal)), col = "gray")
  # Text with the date
  text(x = r * cos(dfinal), y = r * sin(dfinal), labels = dat.France[iend, "date"], srt = 90 - dat.France[iend, "fracYear"]*360, adj = c(0.2, 0.5), cex = 0.75, col = "black", )
  
  dr <- 0.05 * r
  cexT <- 0.8
  text(x = 0, y = r+dr, adj = c(0.5, 0), labels = "2020", col = col2020, cex = cexT)
  if(yearfinal >= 2021) text(x = 0, y = r, adj = c(0.5, 0), labels = "2021", col = col2021, cex = cexT)
  if(yearfinal >= 2022) text(x = 0, y = r-dr, adj = c(0.5, 0), labels = "2022", col = col2022, cex = cexT)
  
  if(addCircle){
  # Add circle to locate the value
    fxy <- polar2xy(valfinal, seq(0, 1, length.out = 100))
    lines(fxy$x, fxy$y, type = "l", col = "black", lty = 1.5, lwd = lwd.grad)
  }
}
```

# Hospitalisations 

```{r}
pxy <- polar2xy(dat.France$hosp, dat.France$fracYear)
```

```{r}
plotHosp <- function(iend = nrow(dat.France), lang = "FR"){
  
  # Initalize plot
  par(pty="s")
  par(mar = rep(3.5, 4))
  rg <- 1.1 * c(- pxy$maxval, pxy$maxval)
  plot(0, 0, type = "n", xlim = rg, ylim = rg, 
       xlab = "", ylab = "", 
       asp = 1, 
       axes = FALSE
  )
  
  par(xpd = TRUE)
  
  graduations1 <- c(10000, 20000, 30000)
  graduations2 <- c(graduations1 - 5000, 35000)
  r <- 40000 # Radius of the month labels
  dfinal <- frac2angle(dat.France[iend, "fracYear"]) # End day of the plot
  valfinal <- dat.France[iend, "hosp"]
  
  addLegendPlot(iend, graduations1, graduations2, r, dfinal, yearfinal =  as.numeric(substr(dat.France[iend, "date"], 1, 4)))
  
  
  if(lang == "FR"){
    title(main = "COVID-19, patients hospitalisés", line = 2.5)
  }else{
    title(main = "COVID-19, hospitalized patients", line = 2.5)
  }
  
  
  #text(x = 0, y = -r - 5000, labels = paste0("Source des données : ", URL, "
  #Mise à jour : ", today, " | ", "@flodebarre"), adj = c(0.5, 1), cex = 0.75)
  
  par(xpd = FALSE)
  
  # Plot values
  points(pxy$x[1:iend], pxy$y[1:iend], col = adjustcolor(cols, 0.8), pch = 16, cex = 0.75)
  
  # Add last points, bigger
  points(pxy$x[iend], pxy$y[iend], col = cols[iend], pch = 16, cex = 2)

  
  #legend(x = "topright", col = c(col2020, col2021), pch = 16, legend = c(expression("2020", col = "blue"), "2021"), cex = 0.8, pt.cex = 0.5, bty = "n")
  
  
  if(lang == "FR"){
    txt <- paste0("@flodebarre, ", today, "
Data : ", URLsource, ", 
Code : https://github.com/flodebarre/covid_indicateurs/blob/main/covid_indicateurs.Rmd"
)
  }else{
    txt <- paste0("@flodebarre, ", today, "
Data: ", URLsource, ", 
Code: https://github.com/flodebarre/covid_indicateurs/blob/main/covid_indicateurs.Rmd"
)
  }
  mtext(side = 1, text = txt, cex = 0.55, line = 2, family = "mono", adj = 0, col = gray(0.4))
  
}
plotHosp(lang  = "EN")

```

```{r}
if(doGif){
  system("rm pics/hosp_*") # Erase previous pictures
  for(iend in 1:nrow(dat.France)){
    png(filename = paste0("pics/hosp_", sprintf("%04d", iend), ".png"), width = 7, height = 7, units = "in", res = 150)
    plotHosp(iend, lang = "EN")
    dev.off()
  }
  
  # Copy the last one to freeze the video
  for(j in 1:10){
    system(paste0("cp pics/hosp_", sprintf("%04d", iend), ".png", " pics/hosp_", sprintf("%04d", iend), "_", j, ".png"))
  }
  
  # Convert into gif (too heavy now)
  #system("convert -quality 100% -delay 0.1 -loop 0 pics/hosp*.png pics/animationHosp.gif")
  
  # Convert into movie (not compatible with Twitter)
  #system("convert -quality 100% -delay 0.12 -loop 0 pics/hosp*.png pics/animationHosp.mp4")
  
  # Convert into movie
  system("rm pics/hosp.mp4")
  system("ffmpeg -pattern_type glob -i 'pics/hosp*.png' -vcodec libx264 -s 700x700 -pix_fmt yuv420p pics/hosp.mp4")
  
  # Remove temporary files
  #system("rm pics/hosp*.png")
  
}

# Just re-do the last one
iend <- nrow(dat.France)
png(filename = paste0("pics/hosp_", sprintf("%04d", iend), ".png"), width = 700, height = 700, pointsize = 20)
plotHosp(iend)
dev.off()

```

# Hospital admissions

```{r}
pxy.adm <- polar2xy(dat.France$incid_hosp_ave7, dat.France$fracYear)
```

```{r}
plotHospAdmissions <- function(iend = nrow(dat.France), lang = "FR"){
  
  # Initalize plot
  par(pty="s")
  par(mar = rep(3.5, 4))
  rg <- 1.1 * c(- pxy.adm$maxval, pxy.adm$maxval)
  plot(0, 0, type = "n", xlim = rg, ylim = rg, 
       xlab = "", ylab = "", 
       asp = 1, 
       axes = FALSE
  )
  
  par(xpd = TRUE)
  
  graduations1 <- c(1000, 2000, 3000)
  graduations2 <- c(graduations1 - 500)
  r <- 3500 # Radius of the month labels
  dfinal <- frac2angle(dat.France[iend, "fracYear"]) # End day of the plot
  valfinal <- dat.France[iend, "incid_hosp_ave7"]
  
  addLegendPlot(iend, graduations1, graduations2, r, dfinal, yearfinal =  as.numeric(substr(dat.France[iend, "date"], 1, 4)))
  
  
  if(lang == "FR"){
    title(main = "COVID-19, nouveaux patients hospitalisés (moyenne glissante sur 7j)", line = 2, font.main = 1)
  }else{
    title(main = "COVID-19, newly hospitalized patients (7-day average)", line = 2, font.main = 1)
  }
  
  
  #text(x = 0, y = -r - 5000, labels = paste0("Source des données : ", URL, "
  #Mise à jour : ", today, " | ", "@flodebarre"), adj = c(0.5, 1), cex = 0.75)
  
  par(xpd = FALSE)
  
  # Plot values
  points(pxy.adm$x[1:iend], pxy.adm$y[1:iend], col = adjustcolor(cols, 0.8), pch = 16, cex = 0.75)
  
  # Add last points, bigger
  points(pxy.adm$x[iend], pxy.adm$y[iend], col = cols[iend], pch = 16, cex = 2)

  
  #legend(x = "topright", col = c(col2020, col2021), pch = 16, legend = c(expression("2020", col = "blue"), "2021"), cex = 0.8, pt.cex = 0.5, bty = "n")
  
  
  if(lang == "FR"){
    txt <- paste0("@flodebarre, ", today, "
Data : ", URLsource, ", 
Code : https://github.com/flodebarre/covid_indicateurs/blob/main/covid_indicateurs.Rmd"
)
  }else{
    txt <- paste0("@flodebarre, ", today, "
Data: ", URLsource, ", 
Code: https://github.com/flodebarre/covid_indicateurs/blob/main/covid_indicateurs.Rmd"
)
  }
  mtext(side = 1, text = txt, cex = 0.55, line = 2, family = "mono", adj = 0, col = gray(0.4))
  
}
plotHospAdmissions(lang  = "EN", iend = nrow(dat.France) - 7)
```

```{r}
if(doGif){
  
  nfinal <- nrow(dat.France) - 3
  system("rm pics/hospAdmin_*") # Erase previous pictures
  for(iend in 1:nfinal){
    png(filename = paste0("pics/hospAdmin_", sprintf("%04d", iend), ".png"), width = 7, height = 7, units = "in", res = 150)
    plotHospAdmissions(iend, lang = "EN")
    dev.off()
  }
  
  # Copy the last one to freeze the video
  for(j in 1:10){
    # As last frame
    system(paste0("cp pics/hospAdmin_", sprintf("%04d", nfinal), ".png", " pics/hospAdmin_", sprintf("%04d", iend), "_", j, ".png"))
  }
  # As first frame
  for(j in 1:5){
    system(paste0("cp pics/hospAdmin_", sprintf("%04d", nfinal), ".png", " pics/hospAdmin_", "0000", "_", j, ".png"))
  }
  

  # Convert into movie
  system("rm pics/hospAdmin.mp4")
  system("ffmpeg -pattern_type glob -i 'pics/hospAdmin*.png' -vcodec libx264 -s 700x700 -pix_fmt yuv420p pics/hospAdmin.mp4")
  
  # Remove temporary files
  #system("rm pics/hospAdmin_*.png")
  
}



```

# Phase plot

Why this animation?
<https://twitter.com/GrahamMedley/status/1503437863622426631?s=20&t=-dWBcPx-7EHTn9s_2ofX4w>

```{r}
MusicSource <- "https://opengameart.org/content/sailor-waltz-with-water-effects-c64-style"
```

```{r}
colW <- met.brewer("Egypt", n = 4, "discrete")[c(1, 2, 4, 3)]

# Wave dates
findDate <- function(itv){
  dat.France[itv, "date"][which(dat.France$incid_hosp_ave7[itv] == min(dat.France$incid_hosp_ave7[itv]))]
}

d1 <- findDate(100:200)
d2 <- findDate(400:500)
d3 <- findDate(570:650)

colsW <- rep(colW[1], nrow(dat.France))
colsW[dat.France$date >= d1] <- colW[2]
colsW[dat.France$date >= d2] <- colW[3]
colsW[dat.France$date >= d3] <- colW[4]
```


```{r}
endLine <- nrow(dat.France) - 4

plotOJHosp <- function(endLine, png = TRUE){
  
  if(png){
    png(paste0("pics/portraitHosp_", sprintf("%04d", endLine), ".png"), 
        width = 6, height = 6.5, res = 150, units = "in")
  }
  ii <- 7:(endLine)
  
  par(las = 1, mgp = c(2.5, 0.5, 0), mar = c(3, 4, 2, 1))
  plot(dat.France$incid_hosp_ave7[ii], dat.France$hosp[ii], col = adjustcolor(colsW[ii], 0.7), pch = 16, 
       xlab = "", ylab = "", 
       xlim = c(0, max(dat.France$incid_hosp_ave7, na.rm = TRUE)), 
       ylim = c(0, max(dat.France$hosp, na.rm = TRUE)))
  
  # Add last point
  points(dat.France$incid_hosp_ave7[endLine], dat.France$hosp[endLine], col = colsW[endLine], pch = 16, cex = 2)
  
  mtext(side = 2, text = "Beds occupied", line = 3, las = 3)
  mtext(side = 1, text = "Daily admissions (7-d average)", line = 1.75)
  title(paste0("France, ", dat.France[ii[1], "date"], "--", dat.France[endLine, "date"]))
  
  if(png){dev.off()}

}

plotOJHosp(nrow(dat.France) - 4)
```

```{r}
if(doGif){
  
  system("rm pics/portraitHosp_*")
  
  for(k in 7:(nrow(dat.France) - 4) ){
    plotOJHosp(k)
  }
  
  # Copy the last one to freeze the video
  for(j in 1:20){
    system(paste0("cp pics/portraitHosp_", sprintf("%04d", k), ".png", " pics/portraitHosp_", sprintf("%04d", k), "_", j, ".png"))
  }
  
 # Convert into movie
  system("rm pics/portraitHosp.mp4")
  system("ffmpeg -pattern_type glob -i 'pics/portraitHosp_*.png' -vcodec libx264 -s 600x650 -pix_fmt yuv420p pics/portraitHosp.mp4")
 
  system("rm pics/portraitHosp_audio.mp4")
  system("ffmpeg -i pics/portraitHosp.mp4 -i data/sailor_waltz_with_water_effects_c64_style.ogg -map 0 -map 1:a -c:v copy -shortest pics/portraitHosp_audio.mp4")
  
}
```


# Rea / soins intensifs

```{r}
pxy <- polar2xy(dat.France$rea, dat.France$fracYear)
```

```{r}
plotRea <- function(iend = nrow(dat.France)){
  
# Initalize plot
par(pty="s")
par(mar = rep(3.5, 4))
rg <- 1.1 * c(- pxy$maxval, pxy$maxval)
plot(0, 0, type = "n", xlim = rg, ylim = rg, 
     xlab = "", ylab = "", 
     asp = 1, 
     axes = FALSE
     )

par(xpd = TRUE)

graduations1 <- c(2000, 4000, 6000)
graduations2 <- c(graduations1 - 1000, 7000)
r <- 8000
dfinal <- frac2angle(dat.France[iend, "fracYear"]) # End day of the plot
valfinal <- dat.France[iend, "rea"]


addLegendPlot(iend, graduations1, graduations2, r, dfinal, valfinal)

if(lang == "FR"){
  title(main = "COVID-19, réanimation ou soins intensifs", line = 2.5)
}else{
  title(main = "COVID-19, intensive care", line = 2.5)
}

#text(x = 0, y = -r - 500, labels = paste0("Source des données : ", URL, "
#Mise à jour : ", today, " | ", "@flodebarre"), adj = c(0.5, 1), cex = 0.75)

par(xpd = FALSE)


points(pxy$x[1:iend], pxy$y[1:iend], col = cols, pch = 16, cex = 0.5)

#legend(x = "topright", col = c(col2020, col2021), pch = 16, legend = c(expression("2020", col = "blue"), "2021"), cex = 0.8, pt.cex = 0.5, bty = "n")

if(lang == "FR"){
  mtext(side = 1, text = paste0("Source des données : ", URLsource, "
Mise à jour : ", today, " | ", "@flodebarre"), cex = 0.75, line = 2)
}else{
  mtext(side = 1, text = paste0("Data : ", URLsource, "
Date : ", today, " | ", "@flodebarre"), cex = 0.75, line = 2)
}

}
plotRea()
```

```{r}
if(doGif){
  for(iend in 1:nrow(dat.France)){
    png(filename = paste0("pics/rea_", sprintf("%04d", iend), ".png"), width = 700, height = 700, pointsize = 20)
    plotRea(iend)
    dev.off()
  }

  # Convert into gif
  system("convert -quality 100% -delay 0.1 -loop 0 pics/rea*.png pics/animationRea.gif")
  
  # Convert into movie
  system("convert -quality 100% -delay 0.12 -loop 0 pics/rea*.png pics/animationRea.mp4")  
  
  # Remove temporary files
  system("rm pics/rea*.png")
}

# Just re-do the last one
iend <- nrow(dat.France)
png(filename = paste0("pics/rea_", sprintf("%04d", iend), ".png"), width = 700, height = 700, pointsize = 20)
plotRea(iend)
dev.off()
```

# Correlation cas / autre indicateur

```{r}
dat.France$pos_ave7 <- sliding.window(dat.France$pos)
dat.France$dc_ave7 <- sliding.window(c(NA, diff(dat.France$dc_tot)))
dat.France$dchosp_ave7 <- sliding.window(c(NA, diff(dat.France$dchosp)))

dat.France$rea_ave7 <- sliding.window(dat.France$incid_rea)


plot(as.Date(dat.France$date), dat.France$pos_ave7, type = "l")
lines(as.Date(dat.France$date), dat.France$incid_hosp_ave7, col = 2)
lines(as.Date(dat.France$date), dat.France$dc_ave7, col = 3)
lines(as.Date(dat.France$date), dat.France$rea_ave7, col = 4)

n <- nrow(dat.France)

corr_hosp <- function(dt){
  v1 <- dat.France[1:(n-dt), "pos_ave7"]
  v2 <- dat.France[(1+dt):n, "incid_hosp_ave7"]
  cr <- mean(v1 * v2, na.rm = TRUE)
  cr
}

corr_mort <- function(dt){
  v1 <- dat.France[1:(n-dt), "pos_ave7"]
  v2 <- dat.France[(1+dt):n, "dc_ave7"]
  cr <- mean(v1 * v2, na.rm = TRUE)
  cr
  }

corr_mort2 <- function(dt){
  v1 <- dat.France[1:(n-dt), "pos_ave7"]
  v2 <- dat.France[(1+dt):n, "dchosp_ave7"]
  cr <- mean(v1 * v2, na.rm = TRUE)
  cr
  }

corr_rea <- function(dt){
  v1 <- dat.France[1:(n-dt), "pos_ave7"]
  v2 <- dat.France[(1+dt):n, "rea_ave7"]
  cr <- mean(v1 * v2, na.rm = TRUE)
  cr
  }

colCas <- rgb(0, 0, 0.75)
colHosp <- rgb(0.9, 0, 0)
colRea <- rgb(0.6, 0, 0.6)
colDeath <- gray(0.4)

dt <- 1
corr_hosp(2)
times <- 0:40
crh <- unlist(lapply(times, corr_hosp))
crd <- unlist(lapply(times, corr_mort))
crd2 <- unlist(lapply(times, corr_mort2))
rea <- unlist(lapply(times, corr_rea))
plot(times, crh/max(crh), ylab = "relative correlation", col = colHosp, xlab = "décalage avec les cas", frame.plot = FALSE)
points(times, crd/max(crd), pch = 2, col = colDeath)
points(times, crd2/max(crd2), pch = 3, col = colDeath)
points(times, rea/max(rea), pch = 3, col = colRea)

argmaxh <- times[which(crh == max(crh))]
argmaxd <- times[which(crd == max(crd))]
argmaxd2 <- times[which(crd2 == max(crd2))]
argmaxrea <- times[which(rea == max(rea))]
abline(v = argmaxh, col = colHosp)
abline(v = argmaxd, col = colDeath)
abline(v = argmaxd2, col = colDeath, lty = 2)
abline(v = argmaxrea, col = colRea, lty = 1)
rm(dt)
c(argmaxh, argmaxd)

xmin <- which(!is.na(dat.France$pos_ave7))[1]
xmin
dat.France2 <- dat.France[xmin:nrow(dat.France), ]

xx <- seq(as.Date("2020-03-01"), as.Date("2021-08-01"), by = "month")

lwdd <- 1.5
ymax <- max(c(dat.France$incid_hosp_ave7/max(dat.France2$incid_hosp_ave7, na.rm = TRUE), dat.France$dc_ave7/max(dat.France2$dc_ave7, na.rm = TRUE)), na.rm = TRUE)

par(las = 1, mgp = c(3, 0.2, 0), tck = -0.01)
plot(as.Date(dat.France$date), dat.France$pos_ave7 / max(dat.France2$pos_ave7, na.rm = TRUE), log = "", col = colCas, frame.plot = FALSE, lwd = lwdd, axes = FALSE, 
     ylim = c(0.01, ymax), 
     type = "n", 
     xlab = "date", ylab = "nombre relatif"#, xlim = range(as.Date(dat.France2$date))
     )

for(i in seq(0, ymax, by = 0.1)){
  abline(h = i, lwd = 0.5, col = gray(0.9))
}

lines(as.Date(dat.France$date), dat.France$pos_ave7 / max(dat.France2$pos_ave7, na.rm = TRUE), lwd = lwdd, col = colCas)


axis(1, at = xx, labels = format.Date(xx, "%Y-%b"), las = 3, cex.axis = 0.7)
axis(2)
lines(as.Date(dat.France[1:(n - argmaxh), "date"]), dat.France[(1+argmaxh):n, "incid_hosp_ave7"]/max(dat.France2[(1+argmaxh:n), "incid_hosp_ave7"], na.rm = TRUE), col = colHosp, lwd = lwdd)
lines(as.Date(dat.France[1:(n - argmaxrea), "date"]), dat.France[(1+argmaxrea):n, "rea_ave7"]/max(dat.France2[(1 + argmaxrea:n), "rea_ave7"], na.rm = TRUE), col = colRea, lwd = lwdd)

#lines(as.Date(dat.France[1:(n - argmaxd), "date"]), dat.France[(1+argmaxd):n, "dc_ave7"]/max(dat.France2[(1+argmaxd:n), "dc_ave7"], na.rm = TRUE), col = colDeath, lwd = lwdd)

lines(as.Date(dat.France[1:(n - argmaxd), "date"]), dat.France[(1+argmaxd):n, "dchosp_ave7"]/max(dat.France2[(1+argmaxd:n), "dchosp_ave7"], na.rm = TRUE), col = colDeath, lwd = lwdd)

add_legend <- function(...) {
  opar <- par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), 
    mar=c(0, 0, 0, 0), new=TRUE)
  on.exit(par(opar))
  plot(0, 0, type='n', bty='n', xaxt='n', yaxt='n')
  legend(...)
}

legend(x = as.Date(dat.France[1, "date"]), y = ymax, legend = c("cas", paste0("nv hospi (J-", argmaxh, ")"), paste0("nv rea (J-", argmaxrea, ")"), paste0("morts hospi (J-", argmaxd2, ")")), col = c(colCas, colHosp, colRea, colDeath), lty = 1, cex = 0.7, inset = c(0, -0.1), xpd = TRUE, yjust = 0, bty = "n", lwd = lwdd, ncol = 2)

```

# By age classes

```{r}
# https://www.data.gouv.fr/en/datasets/donnees-hospitalieres-relatives-a-lepidemie-de-covid-19/
# par classe d'age
URL <- "https://www.data.gouv.fr/en/datasets/r/08c18e08-6780-452d-9b8c-ae244ad529b3"
dataFile <- paste0("data/FranceHosp_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.Hosp <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

head(dat.Hosp)

unique(dat.Hosp$reg)
```


Incidence 
<https://www.data.gouv.fr/en/datasets/donnees-hospitalieres-relatives-a-lepidemie-de-covid-19/>  
`covid-hospit-incid-reg-`

```{r}
URL <- "https://www.data.gouv.fr/en/datasets/r/6fadff46-9efd-4c53-942a-54aca783c30c"

dataFile <- paste0("data/FranceHospIncid_", today, ".csv") # name file with today's date
download.file(URL, dataFile) # download file from repo
dat.HospIncid <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

head(dat.HospIncid)
```


```{r}
# Compute flux from stocks
# Do it first with data for which we have the answer
rm(dat)
dat.0 <- dat.Hosp[dat.Hosp$cl_age90 == 0, ]
unique(dat.0$reg)
dat.cumFrance <- aggregate(dat.0[, -c(1, 3)], by = list(jour = dat.0$jour), FUN = sum)
dat <- dat.cumFrance
```


```{r}
# Merge the datasets into a single one
dat.compare <- merge(dat.France, dat.cumFrance, by.y = "jour", by.x = "date", all = TRUE)

# Checks
any(duplicated(dat.compare$date))
all(dat.compare[order(dat.compare$date), "date"] == dat.compare$date)

# Compare values in the two datasets
plot(as.Date(dat.compare$date), dat.compare$hosp.x - dat.compare$hosp.y)
points(as.Date(dat.compare$date), dat.compare$rea.x - dat.compare$rea.y, col = 2)
points(as.Date(dat.compare$date), dat.compare$rad.x - dat.compare$rad.y, col = 3)
points(as.Date(dat.compare$date), dat.compare$dchosp - dat.compare$dc, col = 4)

```


```{r}
# Compare this to the national indicators
plot(as.Date(dat.cumFrance$jour), dat.cumFrance$hosp, type = "p", ylim = c(0, 35000), lwd = 1, main = "Hospitalisations en cours")
points(as.Date(dat.France$date), dat.France$hosp, col = 2, lty = 2)

plot(as.Date(dat.cumFrance$jour), dat.cumFrance$rea, main = "Rea en cours")
points(as.Date(dat.France$date), dat.France$rea, col = 2, lty = 2)


deltaFlux <- 1 # or 0

# Check 
# Check that no duplicated date
if(any(duplicated(dat$jour))) stop("duplicated dates")

# Check that dates are ordered
if(!all(dat[order(dat$jour), "jour"] == dat$jour)) stop("days need to be ordered")

dat$newrad <- c(NA, diff(dat$rad))
dat$newdc <- c(NA, diff(dat$dc))

if(any(dat$newdc < 0, na.rm = TRUE) | any(dat$newrad < 0, na.rm = TRUE)) stop("negative values")

plot(as.Date(dat$jour), dat$newdc, type = "p", main = "Nouveaux dc")
points(as.Date(dat.France$date), dat.France$incid_dchosp, col = "red")

plot(as.Date(dat$jour), dat$newrad, type = "p", main = "Nouvelles sorties")
points(as.Date(dat.France$date), dat.France$incid_rad, col = "red")

## Calcul incidence hospi
# hosp[t+1] = hosp[t] - rad[t+1] - dc[t+1] + entrees[t+1]
dat$entreeHosp1 <- NA
dat[2:nrow(dat), "entreeHosp1"] <- dat[2:nrow(dat), "hosp"] - dat[1:(nrow(dat)-1), "hosp"] + dat[2:nrow(dat), "newrad"] + dat[2:nrow(dat), "newdc"]

# hosp[t+1] = hosp[t] - rad[t] - dc[t] + entrees[t+1]
dat$entreeHosp0 <- NA
dat[2:nrow(dat), "entreeHosp0"] <- dat[2:nrow(dat), "hosp"] - dat[1:(nrow(dat)-1), "hosp"] + dat[1:(nrow(dat)-1), "newrad"] + dat[1:(nrow(dat)-1), "newdc"]

# hosp[t+1] = hosp[t] - rad[t] - dc[t] + entrees[t]
dat$entreeHosp2 <- NA
dat[1:(nrow(dat)-1), "entreeHosp2"] <- dat[2:nrow(dat), "hosp"] - dat[1:(nrow(dat)-1), "hosp"] + dat[1:(nrow(dat)-1), "newrad"] + dat[1:(nrow(dat)-1), "newdc"]

datt <- merge(dat.France, dat, by.x = "date", by.y = "jour", all = TRUE)
datt[100,]

plot(as.Date(dat.France$date), dat.France$incid_hosp, pch = 16)
points(as.Date(dat$jour), dat$entreeHosp1, col = 2)
points(as.Date(dat$jour), dat$entreeHosp0, col = 3)
points(as.Date(dat$jour), dat$entreeHosp2, col = 4)

plot(datt$incid_hosp, datt$entreeHosp0)
points(datt$incid_hosp, datt$entreeHosp2, col = 3, cex = 0.7)
points(datt$incid_hosp, datt$entreeHosp1, col = 2, cex = 1)
abline(a = 0, b = 1)
legend(x = "topleft", inset = c(0, -0.1), col = c(1, 2, 3), legend = c("calcul 1", "calcul 2", "calcul 3"), pch = c(16, 1, 1), cex = 0.8, horiz = TRUE, yjust = 0, bty = "n")

cor(datt$incid_hosp, datt$entreeHosp1, use =  "pairwise.complete.obs")

```

Define function to compute incidence

```{r}
computeIncidenceHosp <- function(data){
  dat <- data
  
  # Check that no duplicated date
  if(any(duplicated(dat$jour))) stop("duplicated dates")

  # Check that dates are ordered
  if(!all(dat[order(dat$jour), "jour"] == dat$jour)) stop("days need to be ordered")
  
  # The data present the cumulated number of deaths and of 
  # rad; convert into incidence
  dat$newrad <- c(NA, diff(dat$rad))
  dat$newdc <- c(NA, diff(dat$dc))
  
  # Sliding average
  dat$newdc_ave7 <- sliding.window(dat$newdc)
  dat$newrad_ave7 <- sliding.window(dat$newrad)
  dat$hosp_ave7 <- sliding.window(dat$hosp)

  # hosp[t+1] = hosp[t] - rad[t+1] - dc[t+1] + entrees[t+1]
  dat$entreeHosp1 <- NA
  dat[2:nrow(dat), "entreeHosp1"] <- dat[2:nrow(dat), "hosp"] - dat[1:(nrow(dat)-1), "hosp"] + dat[2:nrow(dat), "newrad"] + dat[2:nrow(dat), "newdc"]

  
  # The same on averaged values
  dat$entreesHosp_ave7 <- NA
  dat[2:nrow(dat), "entreesHosp_ave7"] <- dat[2:nrow(dat), "hosp_ave7"] -   dat[1:(nrow(dat)-1), "hosp_ave7"] + dat[1:(nrow(dat)-1), "newrad_ave7"] + dat[1:(nrow(dat)-1), "newdc_ave7"]
  
  dat
}
```

With full dataset

```{r}
dat.merged <- merge(dat.France, computeIncidenceHosp(dat.cumFrance), by.x = "date", by.y = "jour")

par(xpd = FALSE)
plot(dat.merged$incid_hosp, dat.merged$entreesHosp, 
     xlab = "Incidence hosp dans les données indicateurs", 
     ylab = "Incidence hosp calculée", pch = 16, col = rgb(0.5, 0, 0, 0.4), main = "On raw values")
abline(a = 0, b = 1)
text(x = 0, y = 3750, labels = paste("corr =",  signif(cor(dat.merged$incid_hosp, dat.merged$entreesHosp, use = "pairwise.complete.obs"), 4)), adj = c(0, 0.5))

plot(dat.merged$incid_rad, dat.merged$newrad)
# Difference between rad values in the two datasets
plot(dat.merged$rad.x - cumsum(dat.merged$newrad))

plot(as.Date(dat.merged$date), dat.merged$rad.x - dat.merged$rad.y)

points(as.Date(dat.merged$date), dat.merged$dchosp - dat.merged$dc, col = 2)

points(as.Date(dat.merged$date), dat.merged$hosp.x - dat.merged$hosp.y, col = 3)

points(as.Date(dat.merged$date), dat.merged$rea.x - dat.merged$rea.y, col = 4)

# Discordant values
dat.merged[which(dat.merged$rad.x != dat.merged$rad.y | dat.merged$dchosp != dat.merged$dc | dat.merged$hosp.x != dat.merged$hosp.y),]

dat.merged[which(!is.na(dat.merged$tx_incid))[1], "date"]

# -> Recommencer a partir de la
```

There are some issues with the data in the early dates (discordance between the two datasets).  

Do the same starting with later dates, e.g. when case data become available

```{r}
rm(dat.merged)

minCaseDate <- dat.France[which(!is.na(dat.France$tx_incid))[1], "date"]
minCaseDate

dat.merged <- merge(dat.France[dat.France$date >= minCaseDate, ], computeIncidenceHosp(dat.cumFrance[dat.cumFrance$jour >= minCaseDate, ]), by.x = "date", by.y = "jour")

plot(dat.merged$incid_rad, c(diff(dat.merged$rad.x), NA), 
     xlab = "dans données indicateurs", ylab = "dans données par âge", main = "Incidence rad")
points(dat.merged$incid_rad, c(NA, diff(dat.merged$rad.x)), col = 2)
points(dat.merged$incid_rad, dat.merged$newrad, col = 3, pch = 4)

plot(dat.merged$dchosp, dat.merged$dc, xlab = "dc dans données indicateurs", ylab = "dc dans données par âge")
plot(dat.merged$incid_dchosp, c(diff(dat.merged$dchosp), NA), 
     xlab = "dans données indicateurs", ylab = "dans données par âge", main = "Incidence dc")
points(dat.merged$incid_dchosp, c(NA, diff(dat.merged$dchosp)), col = 2)
points(dat.merged$incid_dchosp, dat.merged$newdc, col = 3, pch = 4)

plot(dat.merged$incid_hosp, c(diff(dat.merged$hosp.x), NA) + dat.merged$incid_rad + dat.merged$incid_dchosp)

plot(dat.merged$incid_hosp, c(NA, diff(dat.merged$hosp.x)) + dat.merged$incid_rad + dat.merged$incid_dchosp)
abline(a = 0, b = 1, col = "red", lwd = 2)


plot(dat.merged$incid_hosp, dat.merged$entreeHosp1, xlab = "dans données indicateurs", ylab = "calculée dans données par âge", main = "Incidence hospi")
abline(a = 0, b = 1, col = "cyan")


# Are the dates indeed consecutive? (yes)
unique(diff(as.Date(dat.France$date)))

plot(dat.merged$incid_rad, dat.merged$newrad)
# Difference between rad values in the two datasets
plot(dat.merged$rad.x - cumsum(dat.merged$newrad))

plot(as.Date(dat.merged$date), dat.merged$rad.x - dat.merged$rad.y)

points(as.Date(dat.merged$date), dat.merged$dchosp - dat.merged$dc, col = 2)

points(as.Date(dat.merged$date), dat.merged$hosp.x - dat.merged$hosp.y, col = 3)

points(as.Date(dat.merged$date), dat.merged$rea.x - dat.merged$rea.y, col = 4)

# Discordant values
dat.merged[which(dat.merged$rad.x != dat.merged$rad.y | dat.merged$dchosp != dat.merged$dc | dat.merged$hosp.x != dat.merged$hosp.y),]


```


```{r}
plot(sliding.window(dat.merged$rad.x) - sliding.window(cumsum(dat.merged$newrad)))

# On averaged values
plot(sliding.window(dat.merged$incid_hosp), dat.merged$entreesHosp_ave7, 
     xlab = "Incidence hosp dans les données indicateurs", 
     ylab = "Incidence hosp calculée", pch = 16, col = rgb(0.5, 0, 0, 0.4), main = "On 7-day averaged values")
abline(a = 0, b = 1)

dat.merged$incidH_relativeError <- (dat.merged$incid_hosp - dat.merged$entreesHosp)/dat.merged$incid_hosp
dat.merged$incidH_error <- (dat.merged$incid_hosp - dat.merged$entreesHosp)

plot(dat.merged$incidH_relativeError)
plot(dat.merged$incidH_error)
mean(dat.merged$incid_hosp - dat.merged$entreesHosp, na.rm = TRUE)
range(dat.merged$incid_hosp - dat.merged$entreesHosp, na.rm = TRUE)

```



