---
title: "EPCI data"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
# Load packages
library(MetBrewer) # Colors
library(mapsf) # Maps

for(i in dev.list()) dev.off()
```

# Load data

```{r}
# Whether to download new data
dlData <- FALSE
```

## Epi data

```{r}
# Source: https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-pour-le-depistage-a-compter-du-18-05-2022-si-dep/
# sp-epci-7j-cage65-opendata.csv
URL <- "https://www.data.gouv.fr/fr/datasets/r/26d477f8-4adf-423c-9b17-f54308673c6e"
destFile <- "data/casEPCI.csv"

if(dlData){
  download.file(url = URL, destfile = destFile)
}

# Load data
datEPCI <- read.csv(destFile, sep = ";", dec = ",")
head(datEPCI)

# Clean dates
datEPCI$date1 <- as.Date(substr(datEPCI$sg, 1, 10))
datEPCI$date <- datEPCI$dateMid <- datEPCI$date1 + 3
datEPCI$date2 <- as.Date(substr(datEPCI$sg, 12, 21))

# Dictionary for incidence rate
# (Make sure the values are ordered in increasing order)
tmp <- sort(unique(datEPCI$Ti))[c(1:4, 6:8, 5)]
dicoTi <- 1:length(tmp)
names(dicoTi) <- tmp

dicoTi
# Define the corresponding breaks (check that they match)
breaksTi <- c(0, 10, 20, 50, 150, 250, 500, 1000, Inf)
stopifnot(length(breaksTi) == length(dicoTi) + 1)

# Rewrite the categories in the data
datEPCI$Ti2 <- dicoTi[datEPCI$Ti]


# Dictionary for positivity rate
tmp <- sort(unique(datEPCI$Tp))
dicoTp <- 1:length(tmp)
names(dicoTp) <- tmp
# Rewrite the categories in the data
datEPCI$Tp2 <- dicoTp[datEPCI$Tp]


# Vector of EPCIs
epcis <- unique(datEPCI$epci)

# Vector of dates
dates <- sort(unique(datEPCI$date))
```

```{r, eval = FALSE}
# Explore data
str(datEPCI)

unique(datEPCI$cl_age65)
datEPCI[1,]
```

## Map data

### Mapv1: Ameli

Kept for legacy, but not used: Here we are missing some EPCIs

```{r simplifyMap, eval = FALSE}
# Source map data 
# https://datavaccin-covid.ameli.fr/explore/dataset/georef-france-epci/export/?disjunctive.reg_name&disjunctive.dep_name&disjunctive.epci_name
mapEPCI <- st_read("data/georef-france-epci/georef-france-epci-millesime.shp")

mapEPCI.france <- mapEPCI[as.numeric(substr(mapEPCI$reg_code, 1, 2)) >= 10, ]

# Simplify contours
# Source of the trick
# https://www.r-bloggers.com/2021/03/simplifying-geospatial-features-in-r-with-sf-and-rmapshaper/
library(rmapshaper)
mapEPCI.france_simpl <- ms_simplify(mapEPCI.france, keep = 0.005,
                                keep_shapes = FALSE)

head(mapEPCI)

# Plot the base map
mf_map(x = mapEPCI.france)

mf_map(x = mapEPCI.france_simpl)

save(mapEPCI.france_simpl, file = "data/georef-france-epci/metrosimpler.RData")
```

```{r eval = FALSE}
# Load simplified geographic data
load("data/georef-france-epci/metrosimpler.RData")
# Plot base map
mf_theme(bg = "white")
mf_map(x = mapEPCI.france_simpl)
```

### Mapv2: Banatic

#### Mailand 

I downloaded EPCI shapefiles from the Ministry of Interior. They provide datasets for different years, 
I tried the recent ones, and found that SPF seems to use geographic data from 2019. 

```{r, eval = FALSE}
# eval = FALSE because we only need to do this once

# Source https://www.banatic.interieur.gouv.fr/V5/cartographie/cartographie.php
# Downloaded manually and saved in data/ folder

# I checked 2022, 2021, 2020, EPCIs were missing, 
# and we only have full data for 2019
mapEPCI <- st_read("data/contours_epci/2019/MÇtropole/epcicom2019_region.shp")
# Focus on mainland France
mapEPCI.france <- mapEPCI[as.numeric(mapEPCI$CODE_REG) >= 10, ]

# Simplify shapes to plot the result more quickly
# Source trick
# https://www.r-bloggers.com/2021/03/simplifying-geospatial-features-in-r-with-sf-and-rmapshaper/
library(rmapshaper)
mapEPCI.france_simpl <- ms_simplify(mapEPCI.france, keep = 0.005,
                                keep_shapes = FALSE)

# Plot the base maps to check the simplification
mf_map(x = mapEPCI.france)
mf_map(x = mapEPCI.france_simpl)

# Save the simplified map
save(mapEPCI.france_simpl, file = "data/contours_epci/metro2019simpler.RData")
```

```{r}
# Load simplified contours of EPCIs
load("data/contours_epci/metro2019simpler.RData")
```

#### Outre-Mer

```{r, eval = FALSE}
# Only need to do this once

# Common simplification factor for the maps
simplifyFactor <- 0.05

# Guadeloupe
map.Guadeloupe <- st_read("data/contours_epci/2019/Guadeloupe/contours_epci_971_region.shp")

map.Guadeloupe_simpler <- rmapshaper::ms_simplify(map.Guadeloupe, 
                                    keep = simplifyFactor, 
                                    keep_shapes = FALSE)

mf_map(map.Guadeloupe)
mf_map(map.Guadeloupe_simpler)

# Martinique
map.Martinique <- st_read("data/contours_epci/2019/Martinique/contours_epci_972_region.shp")

map.Martinique_simpler <- rmapshaper::ms_simplify(map.Martinique, 
                                    keep = simplifyFactor, 
                                    keep_shapes = FALSE)
mf_map(map.Martinique)
mf_map(map.Martinique_simpler)

# Guyane
map.Guyane <- st_read("data/contours_epci/2019/Guyane/contours_epci_973_region.shp")

map.Guyane_simpler <- rmapshaper::ms_simplify(map.Guyane, 
                                    keep = simplifyFactor, 
                                    keep_shapes = FALSE)
mf_map(map.Guyane)
mf_map(map.Guyane_simpler)

# Mayotte
map.Mayotte <- st_read("data/contours_epci/2019/Mayotte/Contour_epci_Mayotte_2016_region.shp")

map.Mayotte_simpler <- rmapshaper::ms_simplify(map.Mayotte, 
                                    keep = simplifyFactor, 
                                    keep_shapes = FALSE)
mf_map(map.Mayotte)
mf_map(map.Mayotte_simpler)

# Reunion
map.Reunion <- st_read("data/contours_epci/2019/La RÇunion/contours_epci_974_region.shp")

map.Reunion_simpler <- rmapshaper::ms_simplify(map.Reunion, 
                                    keep = simplifyFactor, 
                                    keep_shapes = FALSE)
mf_map(map.Reunion)
mf_map(map.Reunion_simpler)

save(map.Guadeloupe_simpler, map.Martinique_simpler, map.Guyane_simpler, map.Mayotte_simpler, map.Reunion_simpler, file = "data/contours_epci/outremer2019simpler.RData")
```

```{r}
# Load map data
load("data/contours_epci/outremer2019simpler.RData")
```

## France incidence data

```{r}
# file sp-fra-7j
URL <- "https://www.data.gouv.fr/fr/datasets/r/fd1cf717-1189-4d2a-8b4b-9ad0d2b07a5f"
destFile <- "data/France7j.csv"
if(dlData){
  download.file(URL, destFile)
}

# Read data
datFra <- read.csv(destFile, sep = ";", dec = ",")

# Format date
datFra$date1 <- as.Date(substr(datFra$sg, 1, 10))
datFra$date <- datFra$dateMid <- datFra$date1 + 3
```

# Plot

## Positivity rate

```{r}
max(datEPCI$dateMid)

thedate <- "2022-05-12"
agecl <- 64

# Subset of the data for this combination for date and age class
tmp <- datEPCI[which(datEPCI$date == thedate & datEPCI$cl_age65 == agecl), ]
# Merge with map data
tmp2 <- merge(mapEPCI.france_simpl, tmp, by.y = "epci", by.x = "siren", all.x = TRUE)

# Positivity rate
palTp <- met.brewer("Johnson", n = length(dicoTp), "continuous", direction = -1)
names(palTp) <- dicoTp

uu <- sort(unique(tmp2$Tp2))
mf_theme(bg = "white")
mf_map(tmp2, var = "Tp", type = "typo", pal = palTp[uu], val_order = names(dicoTp)[uu], leg_pos = NA)

mf_legend(type = "typo", pos = "bottomleft", val = rev(names(dicoTp)), pal = rev(palTp), title = "Positivity rate (%)")

mf_title("Positivity rate", bg = "white", fg = "black")
```

## Incidence 

### Mainland 

```{r}
# Categorize incidence France with the same categories as EPCI data
# (will be used for colors)
datFra$TiCat <- cut(datFra$Ti, breaks = breaksTi, right = FALSE) # Categorical version
datFra$TiCatNum <- dicoTi[datFra$TiCat] # Numerical version
```


```{r}
# Define color palette
palTi <- met.brewer("Homer1", n = length(dicoTi), "continuous", direction = -1)
names(palTi) <- dicoTi

# Function to plot Incidence map for a given dataset
plotIncidence <- function(tmp2, legend = TRUE, credits = TRUE, title = TRUE, colBorder = gray(1, 1), lwdBorder = 0.001){
  # tmp2: merged dataset (epi and geo data) to be plotted
  
  # Incidence rate
  uu <- sort(unique(tmp2$Ti2)) # Values of incidence in these data
  
  # Initiate map
  mf_theme(bg = "white")
  
  # Plot as typology map, selecting the colors of the values present in the data
  mf_map(tmp2, var = "Ti", type = "typo", 
         pal = palTi[uu], val_order = names(dicoTi)[uu], 
         leg_pos = NA, border = colBorder, lwd = lwdBorder)
  
  # Add legend
  if(legend) mf_legend(type = "typo", pos = "left", val = rev(names(dicoTi)), pal = rev(palTi), title = "7-d incidence \nper 100'000", val_cex = 0.8, cex = 0.7)
  
  suffix <- ifelse(agecl == 64, " (ages 00-64)", " (ages 65+)")
  
  # Recover date
  dt <- unique(tmp2$date) 
  dt <- dt[!is.na(dt)] # Sometimes there were NAs (when missing EPCIs), remove them

  if(title) mf_title(paste0("7-day incidence, ", dt, suffix), bg = "white", fg = "black")
  
  if(credits) mf_credits(txt = paste0("Data: https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-pour-le-depistage-a-compter-du-18-05-2022-si-dep/\nCode: https://github.com/flodebarre/covid_indicateurs/blob/main/explorationEPCI.Rmd\nMap:  https://www.banatic.interieur.gouv.fr/V5/cartographie/cartographie.php (2019); mapsf ", packageVersion("mapsf")), cex = 0.45, font = 1)
  
}


# Plot map and epi curve

fnI <- paste0("pics/mapEPCI_incidence_", thedate, ".png") # File output name

export <- TRUE # Whether to export the map as png
if(export) mf_export(tmp2, file = fnI, width = 600, height = 600, res = 100)

layout(matrix(1:2, ncol = 1), heights = c(4, 1.5))

# Plot map
plotIncidence(tmp2)

# Plot epi curve, France incidence
# Subset of the data until the `thedate` date
subFra <- datFra[datFra$date <= thedate, ]
par(mar = c(2, 4, 1, 1), mgp = c(2., 0.25, 0), tck = -0.02)
months <- seq(as.Date("2020-01-01"), Sys.Date(), by = "month")
plot(subFra$date, subFra$Ti, col = palTi[subFra$TiCatNum], pch = 16, axes = FALSE, 
     xlab = "", ylab = "", xlim = c(min(dates), max(dates)))
axis(1, at = months, labels = format(months, "%b\n%Y"), cex.axis = 0.8, padj = 0.5)
axis(2, las = 1)
mtext("France", side = 3, adj = 0)

if(export){
  dev.off()
  system(paste("open", fnI))
}

```

```{r}
# Generate maps for all dates
agecl <- 64

system("rm pics/mapEPCI_incidence_*")

for(i in seq_along(dates)){
  # Recover the date
  thedate <- format(as.Date(dates[i], origin = "1970-01-01"), "%Y-%m-%d")
  # Print counter
  print(paste(i, "/", length(dates)))
  # Subset of the data for this combination for date and age class
  tmp <- datEPCI[which(datEPCI$date == thedate & datEPCI$cl_age65 == agecl), ]
  # Merge with geographic data
  tmp2 <- merge(mapEPCI.france_simpl, tmp, by.y = "epci", by.x = "siren", all.x = TRUE)
  
  # Initialize output file
  fnI <- paste0("pics/mapEPCI_incidence_", formatC(i, width = 4, flag = 0), ".png") # Output name
  mf_export(tmp2, file = fnI, width = 600, height = 600, res = 100) # Open device
  
  layout(matrix(1:2, ncol = 1), heights = c(4, 1.5))

  # Plot map
  plotIncidence(tmp2)

  # Plot epi data
  #---
  # Subset of epi data until `thedate` date
  subFra <- datFra[datFra$date <= thedate, ]
  par(mar = c(2, 4, 1, 0.5), mgp = c(1.5, 0.25, 0), tck = -0.025)
  months <- seq(as.Date("2020-01-01"), Sys.Date(), by = "month") # Months for horiz axis
  plot(subFra$date, subFra$Ti, col = palTi[subFra$TiCatNum], pch = 16, axes = FALSE, 
       xlab = "", ylab = "", xlim = c(min(dates), max(dates)))
  axis(1, at = months[months <= thedate], labels = format(months[months<= thedate], "%b\n%Y"), cex.axis = 0.7, padj = 0.5)
  axis(2, las = 1, cex.axis = 0.75)
  mtext("France", side = 3, adj = 0)
  #---
  
  dev.off()
}

# Turn the set of png files into a video
system("rm pics/mapEPCI_incidence.mp4")
system(paste0("ffmpeg -pattern_type glob -i 'pics/mapEPCI_incidence_", "*", ".png' -vcodec libx264 -s 600x600 -pix_fmt yuv420p -filter:v 'setpts=2*PTS' pics/mapEPCI_incidence.mp4"))
system("open pics/mapEPCI_incidence.mp4")
```

### OutreMer

```{r}

thedate <- "2022-05-12"
agecl <- 64

plotOutreMer <- function(thedate = "2022-05-12", agecl = 64){
  # Subset of the data for this combination for date and age class
  tmp <- datEPCI[which(datEPCI$date == thedate & datEPCI$cl_age65 == agecl), ]
  
  layout(matrix(1:6, ncol = 2))

  cB <- gray(0.7, 1) # Color borders
  lB <- 0.5 # lwd borders
  
  # Guadeloupe
  #  Merge epi and geographic data
  tmp2 <- merge(map.Guadeloupe_simpler, tmp, by.y = "epci", by.x = "SIREN_EPCI", all.x = TRUE)
  # Plot typology map
  plotIncidence(tmp2, legend = FALSE, credits = FALSE, title = FALSE, colBorder = cB, lwdBorder = lB)
  # Add title
  mf_title("Guadeloupe", bg = "white", fg = "black")
  
  # Martinique
  tmp2 <- merge(map.Martinique_simpler, tmp, by.y = "epci", by.x = "SIREN_EPCI", all.x = TRUE)
  plotIncidence(tmp2, legend = FALSE, credits = FALSE, title = FALSE, colBorder = cB, lwdBorder = lB)
  mf_title("Martinique", bg = "white", fg = "black")
  
  # Guyane
  tmp2 <- merge(map.Guyane_simpler, tmp, by.y = "epci", by.x = "SIREN_EPCI", all.x = TRUE)
  plotIncidence(tmp2, legend = FALSE, credits = FALSE, title = FALSE, colBorder = cB, lwdBorder = lB)
  mf_title("Guyane", bg = "white", fg = "black")
  
  # Title and legend
  mf_init(tmp2)
  mf_legend(type = "typo", pos = "topleft", val = rev(names(dicoTi)), pal = rev(palTi), title = "7-d incidence per 100'000", val_cex = 0.8, cex = 0.8)
  suffix <- ifelse(agecl == 64, " (ages 00-64)", " (ages 65+)")
  mf_title(paste0("Weekly incidence, ", thedate, ", ages ", suffix), fg = "black", bg = "white")
  mf_credits(txt = paste0("Data: https://www.data.gouv.fr/fr/datasets/donnees-de-laboratoires-pour-le-depistage-a-compter-du-18-05-2022-si-dep/\nCode: https://github.com/flodebarre/covid_indicateurs/blob/main/explorationEPCI.Rmd\nMap:  https://www.banatic.interieur.gouv.fr/V5/cartographie/cartographie.php (2019); mapsf ", packageVersion("mapsf")), cex = 0.45, font = 1)
  
  
  # Mayotte
  tmp2 <- merge(map.Mayotte_simpler, tmp, by.y = "epci", by.x = "siren", all.x = TRUE)
  plotIncidence(tmp2, legend = FALSE, credits = FALSE, title = FALSE, colBorder = cB, lwdBorder = lB)
  mf_title("Mayotte", bg = "white", fg = "black")
  
  # La Réunion 
  tmp2 <- merge(map.Reunion_simpler, tmp, by.y = "epci", by.x = "SIREN_EPCI", all.x = TRUE)
  plotIncidence(tmp2, legend = FALSE, credits = FALSE, title = FALSE, colBorder = cB, lwdBorder = lB)
  mf_title("La Réunion", bg = "white", fg = "black")
}

plotOutreMer()
```

```{r}
system("rm pics/mapEPCI-OutreMer_incidence_*")

for(i in seq_along(dates)){
  # Recover the date
  thedate <- format(as.Date(dates[i], origin = "1970-01-01"), "%Y-%m-%d")
  # Print counter
  print(paste(i, "/", length(dates)))
  
    # Initialize output file
  fnI <- paste0("pics/mapEPCI-OutreMer_incidence_", formatC(i, width = 4, flag = 0), ".png") # Output name
  #png(fnI, width = 600, height = 600, res = 100)
  mf_export(tmp2, file = fnI, width = 600, height = 600, res = 100) # Open device
  
  plotOutreMer(thedate = thedate, agecl = 64)
  
  dev.off()
}


# Turn the set of png files into a video
system("rm pics/mapEPCI-OutreMer_incidence.mp4")
system(paste0("ffmpeg -pattern_type glob -i 'pics/mapEPCI-OutreMer_incidence_", "*", ".png' -vcodec libx264 -s 600x600 -pix_fmt yuv420p -filter:v 'setpts=2*PTS' pics/mapEPCI-OutreMer_incidence.mp4"))
system("open pics/mapEPCI-OutreMer_incidence.mp4")

```

