---
title: "Incidence hospi"
author: "FD"
date: "`r format(Sys.Date(), '%d-%B-%Y')`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(global.par = TRUE)
rm(list = ls())
```


```{r}
today <- Sys.Date() # today's date
```

```{r}
# Whether to generate a gif
doGif <- TRUE
lang <- "EN" #"FR" 

# Whether to download again the data
dlData <- FALSE
```


Mise à jour du `r today`.   
Source des données : <https://www.data.gouv.fr/fr/datasets/synthese-des-indicateurs-de-suivi-de-lepidemie-covid-19/>

# Generic functions and parameters

```{r}
# Function to compute a sliding window 
sliding.window <- function(v, winwdt = 7, pos = 4, na.rm = TRUE){
  # v vector to be averaged/summed
  # winwdt width of the window 
  # pos position of the focal day in the window
  # FUN function to apply
  n <- length(v)
  # Initialize output vector
  out <- 0 * v + (-1)
  out[1:(pos-1)] <- NA
  out[(n + 1 - winwdt + pos) : n] <- NA
  
  for(i in pos : (n - winwdt + pos)){
    out[i] <- mean(v[(i - pos + 1):(i + winwdt - pos)], na.rm = na.rm)
  }
  return(out[1:n])
}
```

```{r}
corr_vv <- function(V1, V2, dt){
  v1 <- v1[1:(n-dt)]
  v2 <- v2[(1+dt):n]
  cr <- mean(v1 * v2, na.rm = TRUE)
  cr
}

```

Dico âges

```{r}

ages <- c("Tous âges", 
                 "0-9",
                 "10-19", 
                 "20-29", 
                 "30-39", 
                 "40-49", 
                 "50-59", 
                 "60-69", 
                 "70-79", 
                 "80-89", 
                 "90+")
names(ages) <- as.character(c(0, seq(9, 89, by = 10), 90))

```

## Graphics

```{r}
colCas <- rgb(0, 0, 0.75)
colHosp <- rgb(0.9, 0, 0)
colRea <- rgb(0.6, 0, 0.6)
colDeath <- gray(0.4)


```

# Load and clean dat


# Load and clean data

## Indicateurs

```{r}
URL <- "https://www.data.gouv.fr/fr/datasets/r/f335f9ea-86e3-4ffa-9684-93c009d5e617"
dataFile <- paste0("data/FranceIndic_", today, ".csv") # name file with today's date
if(dlData){
  download.file(URL, dataFile) # download file from repo
}
dat.France <- read.csv(dataFile, sep = ",", stringsAsFactors = FALSE)
```

Remove leading NAs
```{r}
ibegin <- which(!is.na(dat.France$hosp | !is.na(dat.France$rea)))[1]
dat.France <- dat.France[-(1:ibegin), ]
```

## Hospital data

```{r}
# https://www.data.gouv.fr/en/datasets/donnees-hospitalieres-relatives-a-lepidemie-de-covid-19/
# par classe d'age
URL <- "https://www.data.gouv.fr/en/datasets/r/08c18e08-6780-452d-9b8c-ae244ad529b3"
dataFile <- paste0("data/FranceHosp_", today, ".csv") # name file with today's date
if(dlData){
  download.file(URL, dataFile) # download file from repo
}
dat.Hosp <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)

head(dat.Hosp)
```

Load region names

```{r}
reg_names <- read.csv("data/coderegions.csv")
reg_names

# Change into dictionnary
dic.reg <- reg_names$nom_region
names(dic.reg) <- as.character(reg_names$code_region)

unique(reg_names)
unique(dat.Hosp$reg)
dat.Hosp$region <- dic.reg[as.character(dat.Hosp$reg)]

```


Extract values for France and all age classes

```{r}
# All age classes
dat.0 <- dat.Hosp[dat.Hosp$cl_age90 == 0, ]
# Sum all regions
dat.cumFrance <- aggregate(dat.0[, -c(1, 3, 11)], by = list(jour = dat.0$jour), FUN = sum)
dat <- dat.cumFrance
```

# Compare values

## Compare values in the two datasets

```{r}
# Merge the datasets into a single one
dat.compare <- merge(dat.France, dat.cumFrance, by.y = "jour", by.x = "date", all = TRUE)

# Checks
!any(duplicated(dat.compare$date))
all(dat.compare[order(dat.compare$date), "date"] == dat.compare$date)

# Compare values in the two datasets
plot(as.Date(dat.compare$date), dat.compare$hosp.x - dat.compare$hosp.y, 
     xlab = "Date", ylab = "Indicators - Hospital data")
points(as.Date(dat.compare$date), dat.compare$rea.x - dat.compare$rea.y, col = 2)
points(as.Date(dat.compare$date), dat.compare$rad.x - dat.compare$rad.y, col = 3)
points(as.Date(dat.compare$date), dat.compare$dchosp - dat.compare$dc, col = 4)

# Discordant values
dat.compare[which(dat.compare$rad.x != dat.compare$rad.y | dat.compare$dchosp != dat.compare$dc | dat.compare$hosp.x != dat.compare$hosp.y | dat.compare$rad.x != dat.compare$rad.y),]
```

## Compare incidence values

Cumulative sums 

```{r}
# rad
plot(as.Date(dat.compare$date[-1]), cumsum(dat.compare$incid_rad[-1]), main = "rad", type = "l", xlab = "date", ylab = "value", col = 4)
points(as.Date(dat.compare$date[-1]), dat.compare$rad.x[-1], col = 2, type = "l")
points(as.Date(dat.compare$date[-1]), dat.compare$rad.y[-1], col = 3, type = "l", lty = 2)
legend("topleft", col = c(4, 2, 3), lty = c(1, 1, 2), legend = c("Indicateurs, cumsum incid_rad", "Indicateurs, rad", "Données hosp, rad"), bty = "n")

# dc
plot(as.Date(dat.compare$date[-1]), cumsum(dat.compare$incid_dchosp[-1]), main = "dc hosp", type = "l", xlab = "date", ylab = "value", col = 4)
points(as.Date(dat.compare$date[-1]), dat.compare$dchosp[-1], col = 2, type = "l")
points(as.Date(dat.compare$date[-1]), dat.compare$dc[-1], col = 3, type = "l", lty = 2)
legend("topleft", col = c(4, 2, 3), lty = c(1, 1, 2), legend = c("Indicateurs, cumsum incid_dchosp", "Indicateurs, dchosp", "Données hosp, dc"), bty = "n")

```

Incidence

```{r}
# To find the right formula

# rad
plot(as.Date(dat.compare$date), dat.compare$incid_rad - c(diff(dat.compare$rad.x), NA), main = "rad")
points(as.Date(dat.compare$date), dat.compare$incid_rad - c(NA, diff(dat.compare$rad.x)), col = 2)

dat.compare$newincid_rad <- c(NA, diff(dat.compare$rad.x))

plot(as.Date(dat.compare$date), dat.compare$incid_rad - c(NA, diff(dat.compare$rad.x)), col = 2)

# dc
plot(as.Date(dat.compare$date), dat.compare$incid_dchosp - c(diff(dat.compare$dchosp), NA), main = "rad")
points(as.Date(dat.compare$date), dat.compare$incid_dchosp - c(NA, diff(dat.compare$dchosp)), col = 2)

dat.compare$newincid_dchosp <- c(NA, diff(dat.compare$dchosp))

plot(as.Date(dat.compare$date), dat.compare$incid_dchosp - c(NA, diff(dat.compare$dchosp)), col = 2)
hist(dat.compare$incid_dchosp - c(NA, diff(dat.compare$dchosp)), breaks = seq(-150, 150, by = 1), main = "Comparaison incidences dc", xlab = "Incidence donnée - calculée")

plot(as.Date(dat.compare$date), dat.compare$incid_rad - c(NA, diff(dat.compare$rad.x)), col = 2)
hist(dat.compare$incid_rad - c(NA, diff(dat.compare$rad.x)), breaks = seq(-500, 700, by = 1), main = "Comparaison incidences rad", xlab = "Incidence donnée - calculée")

sum(dat.compare$incid_dchosp, na.rm = TRUE)
dat.compare[nrow(dat.compare), "dchosp"]


# Compute incid_hosp
# hosp[t+1] = hosp[t] - rad[t+1] - dc[t+1] + entrees[t+1]
n <- nrow(dat.compare)
v1 <- c(NA, dat.compare$hosp.x[2:n] - dat.compare$hosp.x[1:(n-1)] + dat.compare$newincid_rad[2:n] + dat.compare$newincid_dchosp[2:n])
plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v1)

v2 <- c(NA, dat.compare$hosp.x[2:n] - dat.compare$hosp.x[1:(n-1)] + dat.compare$newincid_rad[1:(n-1)] + dat.compare$newincid_dchosp[1:(n-1)])
plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v2)
# v1 better

v3 <- c(dat.compare$hosp.x[2:n] - dat.compare$hosp.x[1:(n-1)] + dat.compare$newincid_rad[2:n] + dat.compare$newincid_dchosp[2:n], NA)
plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v3)
# v1 better

v4 <- c(dat.compare$hosp.x[2:n] - dat.compare$hosp.x[1:(n-1)] + dat.compare$newincid_rad[1:(n-1)] + dat.compare$newincid_dchosp[1:(n-1)], NA)
plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v4)

plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v1, col = "red")
points(as.Date(dat.compare$date), dat.compare$incid_hosp - v4, col = "blue")
# v1 still better!

# now v1 with the incidences of the data
v5 <- c(NA, dat.compare$hosp.x[2:n] - dat.compare$hosp.x[1:(n-1)] + dat.compare$incid_rad[2:n] + dat.compare$incid_dchosp[2:n])
plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v5)

plot(as.Date(dat.compare$date), dat.compare$incid_hosp - v1, col = "red")
points(as.Date(dat.compare$date), dat.compare$incid_hosp - v5, col = "cyan")

# Estimation is even better with this formula

```

Summary:   
Incidences for deaths and rad:   
`dat.compare$newincid_rad <- c(NA, diff(dat.compare$rad.x))`

Hosp incidence:  
`c(NA, dat.compare$hosp.x[2:n] - dat.compare$hosp.x[1:(n-1)] + dat.compare$newincid_rad[2:n] + dat.compare$newincid_dchosp[2:n])`

-> Function

```{r}
computeIncidHosp <- function(hosp, cumrad, cumdc){
  # hosp: currently hospitalized
  # cumrand: cumulative number of removed patients
  # cumdc: cumulative number of hospital deaths
  
  # Compute daily new rad and dc
  incid_rad <- c(NA, diff(cumrad))
  incid_dc  <- c(NA, diff(cumdc))
  n <- length(hosp)
  
  # Compute new hosp
  incid_hosp <- c(NA, hosp[2:n] - hosp[1:(n-1)] + incid_rad[2:n] + incid_dc[2:n])
  
  data.frame(incid_hosp = incid_hosp, 
             incid_rad = incid_rad, 
             incid_dc = incid_dc)
}
```

# Incidence data

source <https://legacy.data.gouv.fr/fr/datasets/donnees-relatives-aux-resultats-des-tests-virologiques-covid-19/>

```{r}
URL <- "https://legacy.data.gouv.fr/fr/datasets/r/001aca18-df6a-45c8-89e6-f82d689e6c01"

dataFile <- paste0("data/Cases_", today, ".csv") # name file with today's date
if(dlData){
  download.file(URL, dataFile) # download file from repo
}
dat.Cases <- read.csv(dataFile, sep = ";", stringsAsFactors = FALSE)
head(dat.Cases)

# Add regions
dat.Cases$region <- dic.reg[as.character(dat.Cases$reg)]
```

# By age class now!

```{r}
unique(dat.Hosp$cl_age90)
unique(dat.Cases$cl_age90)

unique(dat.Hosp$reg)
unique(dat.Cases$reg)

reg_names

```

```{r}
for(ag in seq(9, 89, by = 10)){
  fname <- paste0("figFrance_", ag, ".png")
  png(fname, width = 1300, height = 1000, res = 200)
  
  sum(dat.Hosp$cl_age90 == ag)
  sum(dat.Cases$cl_age90 == ag)
  
  # Subselect this age class
  sub.Hosp <- dat.Hosp[which(dat.Hosp$cl_age90 == ag), ]
  sub.Cases <- dat.Cases[which(dat.Cases$cl_age90 == ag), ]
  
  nrow(sub.Hosp)
  
  if(length(unique(sub.Hosp$reg)) > 1){
    # If more than one region
    # Focus on mainland France
    sub.Hosp <- sub.Hosp[which(sub.Hosp$reg >= 10), ]
    sub.Cases <- sub.Cases[which(sub.Cases$reg >= 10), ]
  
    # Add cases of all regions, if multiple regions
    agg.Hosp <- aggregate(sub.Hosp[, -c(1, 2, 3, 11)], by = list(jour = sub.Hosp$jour), FUN = sum)
    
    agg.Cases <- aggregate(sub.Cases[, -c(1, 2, 9, 11)], by = list(jour = sub.Cases$jour), FUN = sum)
    
  }else{
    agg.Hosp <- sub.Hosp
    agg.Cases <- sub.Cases
  }
  
  # Compute incidence
  agg.Hosp[, c("incid_hosp", "incid_rad", "incid_dc")] <- computeIncidHosp(hosp = agg.Hosp$hosp, cumrad = agg.Hosp$rad, cumdc = agg.Hosp$dc)
  
  # Merge datasets
  aggs <- merge(agg.Hosp, agg.Cases, by = "jour")
  
  # 7-day averages
  aggs$P_ave7 <- sliding.window(aggs$P)
  aggs$incid_hosp_ave7 <- sliding.window(aggs$incid_hosp)
  aggs$incid_dc_ave7 <- sliding.window(aggs$incid_dc)
  
  # Find best delay
  # Compute correlation
  times <- 1:35
  n <- nrow(aggs)
  corhosp <- unlist(lapply(times, function(dt) mean(aggs[1:(n-dt), "P_ave7"] * aggs[(1+dt):n, "incid_hosp_ave7"], na.rm = TRUE)))
  argmaxhosp <- times[which(corhosp == max(corhosp))]
  plot(times, corhosp, main = ag)
  abline(v = argmaxhosp)
  
  cordc <- unlist(lapply(times, function(dt) mean(aggs[1:(n-dt), "P_ave7"] * aggs[(1+dt):n, "incid_dc_ave7"], na.rm = TRUE)))
  argmaxdc <- times[which(cordc == max(cordc))]
  plot(times, cordc, main = ag)
  abline(v = argmaxdc)
  
  
  # Plot!
  par(mgp = c(3, 0.25, 0), tck = -0.01, 
      mar = c(5, 3, 3, 3))
  
  lwdd <- 2
  ymax <- 8000
  ymin <- -8000
  alf <- 0.5
  alf.pts <- 0.3
  cex.pts <- 0.4
  pch.pts <- 16
  
  factor <- -10
  
  plot(as.Date(aggs$jour), aggs$P_ave7, 
       col = colCas, lwd = lwdd, 
       type = "l", 
       ylim = c(ymin, ymax), axes = FALSE,
       xlab = "", ylab = "", 
       main = paste0("France, classe d'âge ", ages[as.character(ag)]))
  keep <- !is.na(aggs$P_ave7)
  xx <- as.Date(aggs$jour)[keep]
  yy <- aggs$P_ave7[keep]
  n <- length(xx)
  polygon(x = c(xx, xx[n], xx[1], xx[1]), 
          y = c(yy, 0, 0, yy[1]), 
          col = adjustcolor(colCas, alpha.f = alf), border = adjustcolor(colCas, alpha.f = alf))
  
  points(as.Date(aggs$jour), aggs$P, pch = pch.pts, cex = cex.pts, col = adjustcolor(colCas, alpha.f = alf.pts))
  
  
  xx <- as.Date(aggs$jour)[1:(nrow(aggs)-argmaxhosp)]
  yy <- factor*aggs$incid_hosp_ave7[(1+argmaxhosp):nrow(aggs)]
  lines(xx, yy, col = colHosp, lwd = lwdd)
  keep <- !is.na(yy)
  xx <- xx[keep]
  yy <- yy[keep]
  n <- length(xx)
  polygon(x = c(xx, xx[n], xx[1], xx[1]), 
          y = c(yy, 0, 0, yy[1]), 
          col = adjustcolor(colHosp, alpha.f = alf), border = adjustcolor(colHosp, alpha.f = alf))
  
  points(as.Date(aggs$jour)[1:(nrow(aggs)-argmaxhosp)], factor*aggs$incid_hosp[(1+argmaxhosp):nrow(aggs)], pch = pch.pts, cex = cex.pts, col = adjustcolor(colHosp, alpha.f = alf.pts))
  
  
  
  
  xx <- seq(as.Date("2020-01-01"), as.Date("2021-08-01"), by = "month")
  axis(1, at = xx, labels = format(xx, "%Y-%b"), las = 3, cex.axis = 0.7)
    
  cexx <- 0.9
  axis(2, at = seq(0, ymax, by = 2500), las = 1, col = colCas, col.axis = colCas, cex.axis = cexx)
    
  zz <- seq(0, ymin/factor, by = 250)
  axis(2, at = zz*factor, labels = zz, las = 1, col = colHosp, col.axis = colHosp, cex.axis = cexx)
    
  mtext(side = 3, line = 0, cex = 0.8, text = paste0("lag = ", argmaxhosp, " jours"))
    
  par(xpd = TRUE)
  xxx <- "2020-06-01"
  text(as.Date(xxx), ymax, adj = c(0, 1), labels = "Cas", col = colCas)
  text(as.Date(xxx), ymin, adj = c(0, 0), labels = "Hospitalisations", col = colHosp)
  par(xpd = FALSE)
    
  mtext(paste0(format(Sys.time(), "%Y-%m-%d"), " | @flodebarre, adapté de @jburnmurdoch","
  Données Cas : https://legacy.data.gouv.fr/fr/datasets/donnees-relatives-aux-resultats-des-tests-virologiques-covid-19/  
  Données Hôpitaux : https://www.data.gouv.fr/en/datasets/donnees-hospitalieres-relatives-a-lepidemie-de-covid-19/
  "), col = gray(0.5), side = 1, adj = 1, cex = 0.45, line = 4.25)
  
  rm(xx, xxx, agg.Cases, agg.Hosp, sub.Hosp, sub.Cases, yy)
  
  
  dev.off()
    
  system(paste0("open ", fname))
    
  nchar(fname)
  ff <- substr(fname, 1, nchar(fname) - 4)
  ff
}
  
system("convert -delay 100 figFrance_*.png figFrance.gif")

system("open -a /Applications/Firefox.app figFrance.gif")
```

